---
title: "Lesson 1 - Exercises"
author: "Corné Oosthuizen - OSTAND005"
date: "17 August 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercises

Do the exercises in [Chapter 5](http://r4ds.had.co.nz/transform.html) (data transformation using the **dplyr** verbs) and 
[Chapter 13](http://r4ds.had.co.nz/relational-data.html) (on database joins) of R4DS. 

### Chapter 5

```{r}
suppressWarnings(suppressMessages(library(nycflights13)))
suppressWarnings(suppressMessages(library(tidyverse)))

flights
```


```{r}
sqrt(2) ^ 2
sqrt(2) ^ 2 == 2
#> [1] FALSE
1/49 * 49
1/49 * 49 == 1
#> [1] FALSE
```

```{r}
near(sqrt(2) ^ 2,  2)
#> [1] TRUE
near(1 / 49 * 49, 1)
#> [1] TRUE
```

### 5.2.4 Exercises - `filter()`

1. Find all flights that

* Had an arrival delay of two or more hours
```{r}
filter(flights, arr_delay >= 120)
```

* Flew to Houston (IAH or HOU)
```{r}
filter(flights, dest == 'IAH' | dest == 'HOU')
filter(flights, dest %in% c("IAH", "HOU"))
```
* Were operated by United, American, or Delta
```{r}
levels(as.factor(flights$carrier))
filter(flights, carrier == 'UA' | carrier == 'AA' | carrier == 'DL')
filter(flights, carrier %in% c("UA","AA","DL"))
```
* Departed in summer (July, August, and September)
```{r}
summary(filter(flights, month %in% c(7,8,9))$month)
summary(filter(flights, month == 7 | month == 8 | month == 9)$month)
summary(filter(flights, month >= 7 & month <= 9)$month)
summary(filter(flights, month >= 7, month <= 9)$month)

```

* Arrived more than two hours late, but didn't leave late
```{r}
summary(filter(flights, arr_delay > 120, dep_delay <= 0) %>% select(arr_delay, dep_delay))
```

* Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)
```

* Departed between midnight and 6am (inclusive)
```{r}
summary(select(flights, dep_time))
filter(flights, dep_time == 2400 | dep_time <= 600) %>% arrange(desc(dep_time))
```

2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

* Departed in summer (July, August, and September)
```{r}
summary(filter(flights, month %in% c(7,8,9))$month)
summary(filter(flights, between(month, 6, 9))$month)
```

* Departed between midnight and 6am (inclusive)
```{r}
dim(filter(flights, dep_time == 2400 | dep_time <= 600))
dim(filter(flights, !between(dep_time, 601, 2359)))
```

3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
count( filter(flights, is.na(dep_time)) )
```
```{r}
summary(flights)
```
`dep_time`: Either have not departed yet or will never depart.
`dep_delay`: likewise because the flight did not depart will have no departure delay.
`arr_time`: includes flights that did not depart, and might never have arrived.
`arr_delay`: same as above
`air_time`: same as above

Incomplete data.

4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

```{r}
(NA ^ 0)
```
anything to the power of 0 is 1 - TRUE

```{r}
(NA | TRUE)
```
not available OR TRUE - evaluate to TRUE

```{r}
(FALSE & NA)
```
FALSE & not avialable is FALSE - anything AND FALSE will be FALSE

```{r}
(NA * 0)
```
 `NA` represents an unknown value so missing values are "contagious": almost any operation involving an unknown value will also be unknown.


### 5.3.1 Exercises - `arrange()`

1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).

```{r}
df <- tibble(x = c(5, 2, NA))
arrange(df, desc(is.na(x)))
arrange(df, -(is.na(x)))
```

2. Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
arrange(flights, desc(dep_delay))
arrange(flights, dep_delay)
```

3. Sort flights to find the fastest flights.
```{r}
arrange( flights, air_time)
```

4. Which flights travelled the longest? Which travelled the shortest
```{r}
arrange(flights, desc(distance)) %>% select(tailnum, distance)
arrange(flights, distance) %>% select(tailnum, distance)
```

### 5.4.1 Exercises - `select()`

There are a number of helper functions you can use within select():

* starts_with("abc"): matches names that begin with "abc". (starts with a prefix - `starts_with(match, ignore.case = TRUE, vars = current_vars())`)

* ends_with("xyz"): matches names that end with "xyz". (ends with a prefix - `ends_with(match, ignore.case = TRUE, vars = current_vars())`)

* contains("ijk"): matches names that contain "ijk". (contains a literal string - `contains(match, ignore.case = TRUE, vars = current_vars())`)

* matches("(.)\\1"): selects variables that match a regular expression. This one matches any variables that contain repeated characters. You'll learn more about regular expressions in strings. (matches a regular expression - `matches(match, ignore.case = TRUE, vars = current_vars())`)

* num_range("x", 1:3) matches x1, x2 and x3. (a numerical range like x01, x02, x03 - `num_range(prefix, range, width = NULL, vars = current_vars())`)

* one_of(): variables in character vector.

* everything(): all variables.


1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.

```{r}
dim( select(flights, dep_time, dep_delay, arr_time, arr_delay) )
dim( select(flights, starts_with("dep"), starts_with("arr")) )
dim( select(flights, matches("^(dep|arr)")) )
```


2. What happens if you include the name of a variable multiple times in a select() call?
```{r}
select(flights, dep_time, dep_time)
```

Only get the column once.

3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")

select(flights, vars)
select(flights, one_of(vars))
```

`one_of()`: variables in character vector - so the list of column names returned.

4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
select(flights, contains("TIME"))
```

No, `contains(match, ignore.case = TRUE, vars = current_vars())` by default ignores the case of the match string, this can be changed by overwriting the `ignore.case` variable.

```{r}
select(flights, contains("TIME", ignore.case = FALSE))
```

### 5.5.2 Exercises - `mutate()` and `transmute()`

1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they're not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
select(flights, dep_time, sched_dep_time) %>%
  mutate( dep_time_st = str_c(str_sub(dep_time,1,-3), str_sub(dep_time,-2,-1), sep = ":"),
          sched_dep_time_st = str_c(str_sub(sched_dep_time,1,-3), str_sub(sched_dep_time,-2,-1), sep = ":"),
          dep_time_min = (dep_time %/% 100)*60 + (dep_time %% 100), 
          sched_dep_time_min = (sched_dep_time %/% 100)*60 + (sched_dep_time %% 100))
```

2. Compare `air_time` with `arr_time - dep_time`. What do you expect to see? What do you see? What do you need to do to fix it?
```{r}
select(flights, dep_time, arr_time, air_time) %>%
  mutate( dep_time_min = (dep_time %/% 100)*60 + (dep_time %% 100),
          arr_time_min = (arr_time %/% 100)*60 + (arr_time %% 100),
          real_air_time = (arr_time_min - dep_time_min) %% (60*24),
          flight_diff = real_air_time - air_time) %>%
  arrange(real_air_time)
```
* Well we expect to see a numerical representation of the flight time that matches `air_time`, we get a difference and in some cases a alrge negative value.
* The negative value is due to flights that start before midnight and land after (next day) so the `real_air_time` should take into account that a day might have passed.
* There is still some inaccuracy between `air_time` and `real_air_time` and might be due to the measurement of what `air_time` is, actual flight time or take-off and landing included.

3. Compare `dep_time`, `sched_dep_time`, and `dep_delay`. How would you expect those three numbers to be related?

```{r}

select(flights, dep_time, sched_dep_time, dep_delay) %>%
  mutate( dep_time_min = (dep_time %/% 100)*60 + (dep_time %% 100),
          sched_dep_time_min = (sched_dep_time %/% 100)*60 + (sched_dep_time %% 100)) %>%
  #arrange(desc(real_dep_delay))
  transmute(real_dep = near((sched_dep_time_min + dep_delay) %% (60*24), dep_time_min, tol=1)) %>% # take into account next day and tolerance of 1
  summary()  

```
`sched_dep_time` + `dep_delay` = `dep_time`

So mostly the scheduled and departure times match up.

4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

What does 1:3 + 1:10 return? Why?

What trigonometric functions does R provide?


1. Refer back to the lists of useful **mutate** and **filtering** functions. Describe how each operation changes when you combine it with grouping.

`group_by()` changes the scope of each function from operating on the entire dataset to operating on it group-by-group.




2. Which plane (tailnum) has the worst on-time record?

What time of day should you fly if you want to avoid delays as much as possible?

For each destination, compute the total minutes of delay. For each, flight, compute the proportion of the total delay for its destination.

Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using lag() explore how the delay of a flight is related to the delay of the immediately preceding flight.

Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?

Find all destinations that are flown by at least two carriers. Use that information to rank the carriers.

For each plane, count the number of flights before the first delay of greater than 1 hour.

