---
title: "Lesson 2 - Exercises"
author: "Corné Oosthuizen - OSTAND005"
date: "21 August 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=TRUE, eval = TRUE}
suppressMessages(library(tidyverse))
```

## Exercises

1. Adapt the pairwise similarity function so that it doesn't use loops.

```{r}

# Exercises:
# 1. Adapt the pairwise similarity function so that it doesn't use loops.

# cosine_sim <- function(a,b){ crossprod(a,b) / sqrt(crossprod(a)*crossprod(b)) }

cosineDist <- function(x) {
  as.matrix( tcrossprod(x) / sqrt(rowSums(x^2) %*% t(rowSums(x^2))) ) 
}

b <- cosineDist(viewed_movies)
row.names(b) <- row.names(viewed_movies)
colnames(b) <- row.names(viewed_movies)
#b
```

2. Implement a k-nearest-neighbours version of item-based collaborative filtering.

```{r}
library(caret)

load("output/recommender.RData")
```

```{r}
set.seed(2)

knn.df   <- viewed_movies[sample(nrow(viewed_movies)),]
knn.df$userId <- as.factor(knn.df$userId)

train.df <- knn.df[1:as.integer(0.7*30),]
test.df  <- knn.df[as.integer(0.7*30 +1):30,]

control <- trainControl(method = "cv", number = 10)
metric <- "Accuracy"

fit.knn <- train(userId ~ ., data = viewed_movies, method = "knn", metric = "RMSE", trControl = control)
```




data.train <- cbind(rownames(viewed_movies), as.data.frame(viewed_movies))
colnames(data.train) <- paste(rep("m",ncol(data.train)),c(1:ncol(data.train)),sep="")
names(data.train)[1] = "users"

trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

knn_fit <- train(users ~ . , data = data.train, method = "knn",
 trControl=trctrl,
 preProcess = c("center", "scale"),
 tuneLength = 10)

```

3. Adapt the `recommender_accuracy()` function so that it can be used with an arbitrary number of users and movies.


4. Experiment with the optimizers used in the matrix factorization collaborative filter.



