---
title: "Data Science for Industry - Project 2"
author: "Corné Oosthuizen - OSTAND005"
date: "16 September 2017"
output: 
  html_document:
    toc: yes
    toc_depth: 2
  comment: '>'
  word_document: default
---
<style>
   h1 { font-size: 28px; }
   h2 { font-size: 24px; }
   h4 { margin-bottom: 0; color: #193441;}
   table h4 { margin-bottom: 0; padding-left: 1em;}
   table table { margin: 0 1em; width: auto; max-width: auto; }
   #TOC ul { list-style: none; }
   #TOC ul ul { list-style: square; }

   h4.date {
    font-size: 16px;
    color: #888;
    margin-bottom: 1em;
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE)
options(scipen=1, digits=3)
options(repr.plot.width=4, repr.plot.height=3) # set plot size in the notebook
```

```{r init}
library(tidyverse)
library(tidytext)
library(stringr)
library(lubridate)
library(formattable)

# Replace specified values with new values, in a factor or character vector.
revalue <- plyr::revalue

# Load the data
load("../data/complaints.RData")

complaints <- complaints %>% 
  mutate(date = parse_date(date_received, "%m/%d/%Y")) %>%
  mutate(month = make_date(year(date), month(date)))

# Remove unwanted characters: 
# - Brackets
# - XXXX
# - ellipses, hyphens, slashes
# - puncuation
# - urls
replace_reg <- "\\n|\\(|\\)|\\{|\\}|(XXXX)|(XXX)|(XX)|(X)|RE|&|&amp;|&lt;|&gt;|'s|n\\/a|\\$?([0-9])+|(_+)|-|\\%|\\/|\\.\\.\\.|\177|\032|#|\\@|;|'|,|\\.|\"|http(.*)(.html|.htm|.aspx|.php)"
unnest_reg <- "([^A-Za-z_\\d#@']|'(?![A-Za-z_\\d#@]))"

# replace contractions
complaints$consumer_complaint_narrative %<>% 
      gsub("'m", "am", .) %>%
      gsub("'re", "are", .) %>%
      gsub("'ll", "will", .) %>%
      gsub("'d", "had", .) %>%
      gsub("n't", "not", .) %>%
      gsub("wo n't", "would not", .) %>%
      gsub("ca n't", "cannot", .) %>%
      gsub(pattern = "\\b\\w{16,}", replacement = "") %>% # remove strings longer than 16 characters
      gsub(pattern = "(\\w)\\1{2,}", replacement = "\\1") # replace repeating characters 2 or more with single one

complaints.tidy <- complaints %>%
  mutate(text = str_replace_all(consumer_complaint_narrative, replace_reg, " "), compensated = consumer_compensated) %>%
  unnest_tokens(word, text, token = "regex", pattern = unnest_reg, to_lower = TRUE) # tokenize
    
complaints.no_words <- complaints.tidy %>% group_by(month, id) %>% count(word) %>% summarise(n = sum(n))

complaints.tidy %<>% 
  anti_join( filter(stop_words, lexicon == "snowball"), by = c("word" = "word")) %>% # remove stop words
  filter(word != "x") %>%
  select(month, word, compensated, product, id)      # choose the variables we need

save(complaints.tidy, complaints.no_words, complaints, file = "../data/complaints.tidy.RData")
```
Each record/row captures a single complaint, on the following variables:

Field | Description
------------- | -------------
product | a broad categorisation of the type of financial product involved (5 categories: bank account or service, credit card, credit reporting, debt collection, mortgage)
consumer_complaint_narrative | the submitted text explaining the nature of the complaint.
consumer_compensated | a binary indicator of whether the consumer was compensated (monetarily or non-monetarily) after the complaint.
date_received | the date the complaint was received

```{r}
select(complaints, id, product, date, consumer_compensated, consumer_complaint_narrative) %>% 
  head() %>%
  formattable(row.names = FALSE)
```


```{r}
complaints.compensated <- complaints %>% group_by(consumer_compensated) %>% summarise(count = n())
complaints.compensated
```

```{r}
complaints.products <- complaints %>% group_by(product) %>% summarise(count = n())
complaints.products
```

a.	Use the bing sentiment dictionary to calculate sentiment scores for the complaints.
```{r}

pallete.2 <- c("#E97E3B", "#3984B6")
pallete.3 <- c("#E97E3B", "#85B942", "#3984B6")
pallete.4 <- c("#CF4F29", "#F3B355", "#264992", "#5ABAD1")
pallete.6 <- c("#CF4F29", "#F3B355", "#85B942", "#B0DE63", "#264992", "#5ABAD1")

complaints.bing <- complaints.tidy %>% 
  left_join(get_sentiments("bing")) %>% # add sentiments (pos or neg)
  select(word, sentiment, everything()) %>%
  mutate(sentiment = ifelse(is.na(sentiment), "neutral", sentiment))

complaints.bing %>%
  filter(sentiment == "positive") %>% # only positive
  count(word) %>%
  top_n(20, wt = n) %>% 
  ggplot(aes(reorder(word,n),n)) + geom_col(fill="#5ABAD1", colour="#264992") + coord_flip() + xlab("")
```
```{r}
complaints.bing %>%
  filter(sentiment == "negative") %>%
  count(word) %>%
  top_n(20, wt = n) %>% 
  ggplot(aes(reorder(word,n),n)) + geom_col(fill="#F3B355", colour="#CF4F29") + coord_flip() + xlab("")
```

b.	Plot the histogram of sentiment scores.

```{r}
# sentiments per month all of them
sentiments.all <- complaints.bing %>%
  group_by(month, sentiment) %>%
  summarize(n = n()) 

sentiments.all %>%
  ggplot(aes(x = month, y = n, fill = sentiment)) +
    geom_bar(stat="identity", position = "dodge") +
    scale_fill_manual(values = pallete.3)
    

sentiments.all %>%
  filter(sentiment != "neutral") %>%
  ggplot(aes(x = month, y = n, fill = sentiment)) +
    geom_bar(stat="identity", position = "dodge") +
    scale_fill_manual(values = pallete.2)  
```

```{r}

# sentiments per month byt whether a cosumer was compensated
sentiments.compensated <- complaints.bing %>%
  mutate(type = paste0(sentiment, " ", compensated)) %>%
  group_by(month, sentiment, type) %>%
  summarize(n = n())

sentiments.compensated %>%
  ggplot(aes(x = month, y = n, fill = type)) +
    geom_bar(stat="identity", position = "dodge") +
    scale_fill_manual(values = pallete.6)

sentiments.compensated %>%
  filter(sentiment != "neutral") %>%
  ggplot(aes(x = month, y = n, fill = type, label = n)) +
    geom_bar(stat="identity", position = "dodge", width = 40) +
    scale_fill_manual(values = pallete.4) + 
    geom_text(size = 3, position = position_stack(vjust = 0.5), family = "Times New Roman")
```
```{r}
# sentiments per month byt whether a cosumer was compensated
# sentiments.compensated <- complaints.bing %>%
#   mutate(type = paste0(sentiment, " ", compensated)) %>%
#   group_by(month, sentiment, type) %>%
#   summarize(n = n())

complaints.false <- complaints.compensated %>% filter(consumer_compensated == FALSE) %>% select(count) %>% as.numeric()
complaints.true <- complaints.compensated %>% filter(consumer_compensated == TRUE) %>% select(count) %>% as.numeric()

sentiments.normalized <- sentiments.compensated %>%
  mutate( n = ifelse( grepl("FALSE",type), n / complaints.false, n / complaints.true))

sentiments.normalized %>%
  filter(sentiment != "neutral") %>%
  ggplot(aes(x = month, y = n, fill = type)) +
    geom_bar(stat="identity", position = "dodge") +
    scale_fill_manual(values = pallete.4)

```
```{r, message=TRUE}

# sentiments per month byt whether a cosumer was compensated
sentiments.product <- complaints.bing %>%
  mutate(type = paste0(sentiment, " ", product)) %>%
  group_by(month, sentiment, type) %>%
  summarize(n = n())


for (i in 1:nrow(sentiments.product)) {
  
  sentiments.product[i,]$n = sentiments.product[i,]$n / complaints.products[complaints.products$product == str_replace(sentiments.product[i,]$type,"(negative )|(positive )|(neutral )",""),]$count
}

sentiments.product %>%
  filter(sentiment == "negative") %>%
  ggplot(aes(x = month, y = n, fill = type)) +
    geom_bar(stat="identity", position = "dodge") +
    scale_fill_manual(values = c("#FDEA6F","#F3B355","#E97E3B","#CF4F29","#723C2C"))

sentiments.product %>%
  filter(sentiment != "neutral") %>%
  ggplot(aes(x = month, y = n, fill = type)) +
    geom_bar(stat="identity", position = "dodge") + 
    scale_fill_manual(values = c("#B7DFCB","#5ABAD1","#3984B6","#264992","#161F63"))
```


```{r}


# sentiments per month byt whether a cosumer was compensated
test <- complaints.bing %>% group_by(month) %>% count(sentiment, product, compensated)


test %>%
  filter(sentiment != "negative") %>%
  ggplot(aes(x = month, y = n, group = interaction(product, compensated), fill = interaction(product, compensated))) +
    geom_bar(stat="identity", position = "dodge") +
    scale_fill_discrete(name="Product + Compensated")
```


c.	Do topic modelling using LDA, The user should be able to specify how many topics they want to use (say between 2 and 5 topics).

d.	Extract and plot the word-topic probabilities (i.e. “beta”) for the 15 most popular words in each topic. 



a.	Return the sentiment score of the new complaint
b.	Return the quantile of the sentiment score of the new complaint (i.e. the relative rank compared to other complaints)
c.	Return the topic probabilities for the new complaint (i.e. “gamma”: see the use of the “posterior” function given in the help file for the LDA function “?LDA”). 


